---
title: "Result Analysis"
---

First, we will clean the results so that we can have everything better organized. 
These are the results of the multiQTL experiment, with different models. These were obtained
without the Q models. 

```{r,eval=F}
files<-rep(paste0("Storage/ Multi_",c("LM","MM")),each=3)
files<-paste0(files,"_",c("biallelic","ancestral","parental")," .RDS")
names<-c("mLB","mLA","mLP","mMB","mMA","mMP")

for(i in 1:length(files)){#with this loop we create a file with all QTLs for each situation
  res<-readRDS(files[i])
  res1<-lapply(res,function(x) x[[1]])
  
  if(any(names[i]==c("mMB","mMA","mMP"))){
    res1<-lapply(res1,t)
  }
  
  #extract each of the heratibility experiment values
  res02<-sapply(res1,function(x) x[1,])
  res05<-sapply(res1,function(x) x[2,])
  res08<-sapply(res1,function(x) x[3,])
  colnames(res02)<-names(res)
  colnames(res05)<-names(res)
  colnames(res08)<-names(res)
  
  #with this we create a variable for each experiment
  assign(names[i],list(h02=res02,h05=res05,h08=res08))
}

h02<-lapply(list(mLB,mLA,mLP,mMB,mMA,mMP),function(x) x[[1]])
h05<-lapply(list(mLB,mLA,mLP,mMB,mMA,mMP),function(x) x[[2]])
h08<-lapply(list(mLB,mLA,mLP,mMB,mMA,mMP),function(x) x[[3]])

names(h02)<-names
names(h05)<-names
names(h08)<-names

Exp2<-list(h02,h05,h08)
names(Exp2)<-c("h02","h05","h08")
saveRDS(Exp2,"Storage/Experiment2.RDS")

rm("mLB","mLA","mLP","mMB","mMA","mMP","h02","h05","h08")
```

Here, the processed outcome is used to generate comparison plots. The study of these p-values led to the conclusion that there was some issues with linear model p-value calculation. Which lead to experiments 3 and 4 (repeats of experiment 2 and 1 respectively)
```{r}
source("mpQTL_fun.R")
Exp2<-readRDS("Storage/Experiment2.RDS")
a<-7e-5 #LiJi correction yields systematically around this value

#we obtain with this the object pos, which contains the absolute position, in cM, for each marker
map<-data.table::fread("PedigreeSIM/Potato.map")
cumchrom<-cumsum(sapply(unique(map$chromosome),function(x)
  max(map$position[which(map$chromosome==x)])))
cumchrom<-c(0,cumchrom)
pos<-map$position+cumchrom[map$chromosome]
rm(cumchrom,map)

#With this we generate the plots for Experiment 2
limits<-list(NAM3=1:25,NAM7=26:50,NAM10=51:75)
QTLs<-list(NAM3=c(6906,3122,930),
           NAM7=c(6906,3122,930,60,5727,3884,1794),
           NAM10=c(6906,3122,930,60,5727,3884,1794,6492,5239,4856))

herits<-names(Exp2) #the names of heritabilities
models<-names(Exp2[[1]]) #the names of the models
ancs<-names(limits) #the names of the ancestral groups

### Model comparison plots ###
for(i in 1:length(Exp2)){#each of the different heritabilities
  h2<-herits[i]
  
  for(j in 1:length(limits)){# for each NAM pop
    anc<-ancs[j]
    QTL<-QTLs[[j]]
    
    png(paste0(paste0("Storage/Plots/",anc,"_",h2,"BIG.png")),width=2000,height=1200)
    comp.skyplot(Exp2[[i]],pos,
                 index=limits[[j]],QTLs=pos[QTLs[[j]]],
                 main=paste0("Model comparison ",h2," ",anc),
                 c=80,h=c(140,250),coltype = "divergent")
    dev.off()
    
    png(paste0(paste0("Storage/Plots/","Mixed",anc,"_",h2,".png")),width=1000,height=600)
    comp.skyplot(Exp2[[i]][4:6],pos,
                 index=limits[[j]],QTLs=pos[QTLs[[j]]],
                 main=paste0("Mixed model comparison ",h2," ",anc),
                 c=80,h=c(250),coltype = "sequential")
    dev.off()
    
    png(paste0(paste0("Storage/Plots/","Linear",anc,"_",h2,".png")),width=1000,height=600)
    comp.skyplot(Exp2[[i]][1:3],pos,
                 index=limits[[j]],QTLs=pos[QTLs[[j]]],
                 main=paste0("Linear model comparison ",h2," ",anc),
                 c=80,h=c(140),coltype = "sequential")
    dev.off()

  }
  
}

### Single model plots ###
for(i in 1:length(Exp2)){#each of the different heritabilities
  h2<-herits[i]
  
  for(j in 1:length(limits)){# for each NAM pop
    anc<-ancs[j]
    QTL<-QTLs[[j]]
    
    png(paste0(paste0("Storage/Plots/",anc,"_",h2,".png")),width=1000,height=600)
    comp.skyplot(Exp2[[i]],pos,index=limits[[j]],QTLs=pos[QTLs[[j]]],main=paste0("Model comparison ",h2))
    dev.off()
    
    for(k in 1:length(Exp2[[i]])){# for each model
      mod<-models[k]
      result<-Exp2[[i]][[k]][,limits[[j]]]
      plotname<-paste0("Storage/Plots/",anc,"_",h2,"_",mod)
      name<-paste(anc,h2,mod)
      
      #QTL cummulative plot
      png(paste0(plotname,"_skyline.png"),width=1200,height=800,antialias = 50)
      colplotQTL(result,pos,alpha=a,main=paste0(name," QTL plot"),
                 QTLs=pos[QTL])
      dev.off()
      
      #Superposition of QQplots
      png(paste0(plotname,"_superQQ.png"),width = 800,height = 800)
      colQQplot(result,main=paste(name,"all QQ-plots"))
      dev.off()
      
      #QQplots 
      png(paste0(plotname,"_QQplot.png"),width = 800,height = 800)
      ggd.qqplot(result,main=paste(name,"total QQ-plot"))
      dev.off()
      
      #pvalue histogram
      png(paste0(plotname,"_pvalhist.png"),width = 800, height = 800)
      hist(result,main=paste(name,"pvalue histogram"))
      dev.off()
    }    

  }
  
}

```

By calculating LiJi threshold values for the different NAMs, we saw that all thresholds were around 7.8e-5. The computation can take a long time, as a full marker correlation matrix must be calculated.
```{r}
#Calculate the threshold for each experiment
dosfiles<-paste0("PedigreeSIM/NAM_crosses/",c("1_","3_","7_","10_"),
                "ancestral/")
dosfiles<-paste0(dosfiles,"cross",sprintf("%03.f",c(0,200,600,900)))
dosfiles<-paste0(dosfiles,"_alleledose.dat")
map<-data.table::fread("PedigreeSIM/Potato.map")


alpha<-sapply(dosfiles,function(file){
  
  dosage<-data.table::fread(file,header=T)
  dosage<-as.matrix(dosage[,-1])
  segregants<-apply(dosage,1,function(x) length(unique(x))!=1)
  
  a<-LiJi(dosage[segregants,],0.05)
  return(a)
})



cor(t(dosage))
```

More results for experiment 2
```{r}

for(i in 1:length(herits)){#each of the different heritabilities
  h2<-herits[i]
  
  for(j in 3){# for each NAM pop
    anc<-ancs[j]
    QTL<-QTLs[[j]]
    
    bmp(paste0(paste0("Storage/Plots/",anc,"_",h2,".bmp")),width=800,height=600,res=115)
    comp.skyplot(lapply(Exp2[[i]][4:6],function(x) x[-6910:-6890,])
                 ,pos[-6910:-6890],
                 index=limits[[j]], threshold=7e-100,
                 main=paste0("Mixed model comparison ",h2," ",anc),
                 c=80,h=c(250),coltype = "sequential",min=F)
    dev.off()

  }
}

```

to obtain the "pheno" objects that are used in these plots, the phenotyping function must be run manually, a pheno object is included in the function, which contains genetic effects for each locus and allows to better identify genetic effects.
```{r}
  png("Storage/Plots/NAM3_Geneffect.png",width =600,height=500)
  plot(c(rep(0,10),rep(1:9,each=50))+rnorm(460,sd=0.05),rowSums(pheno),
       main="Genetic components NAM3",xlab="Cross",ylab="Genetic effect",
       ylim=c(0,1.5))
  abline(v=c(0.5,2.5,5.5))
  text(c(0,1.5,4,7.5),rep(1.5,4),c("parents","A1xA1","A1xA2","A1xA3"))
  dev.off()
  
  png("Storage/Plots/NAM3_Total.png",width =600,height=500)
  plot(c(rep(0,10),rep(1:9,each=50))+rnorm(460,sd=0.05),result,
       main="Phenotypes NAM3",xlab="Cross",ylab="Phenotype",
       ylim=c(min(result),max(result)+1))
  abline(v=c(0.5,2.5,5.5))
  text(c(0,1.5,4,7.5),rep(max(result)+1,4),c("parents","A1xA1","A1xA2","A1xA3"))
  dev.off()
  
  data<-data.frame(pop=c(rep("parent",10),rep(1:9,each=50)),pheno=result)
  boxes<-list(data$pheno[which(data$pop%in%c(1,2))],
         data$pheno[which(data$pop%in%c(3,4,5))],
         data$pheno[which(data$pop%in%c(6,7,8,9))])
 
  png("Storage/Plots/NAM3_Ancestral_cross.png",width=600,height=500)
  names(boxes)<-c("A1xA1","A1xA2","A1xA3")
  boxplot(boxes,main="Boxplot of ancestral crosses",xlab="Cross",ylab="Phenotype")
  dev.off()
  
  png("Storage/Plots/NAM3_Boxplot_cross.png",width=600,height=500)
  boxplot(pheno~pop,data=data,main="Boxplot NAM3 per cross",
          xlab="Cross",ylab="Phenotype")
  abline(v=c(2.5,5.5,9.5))
  text(c(1.5,4,7.5),rep(66,4),c("A1xA1","A1xA2","A1xA3"))
  dev.off()
  

  
  data<-data$pop[order(result)]
  ##### NAM10
  
    png("Storage/Plots/NAM10_Geneffect.png",width =600,height=500)
  plot(c(rep(0,10),rep(1:9,each=50))+rnorm(460,sd=0.05),rowSums(pheno),
       main="Genetic components NAM10",xlab="Cross",ylab="Genetic effect",
       ylim=c(0,max(rowSums(pheno))+2))
  abline(v=c(0.5:8.5))
  text(c(0:10),rep(40,10),c("parents",paste0("A0xA",1:9)))
  dev.off()
  
  png("Storage/Plots/NAM10_Total.png",width =600,height=500)
  plot(c(rep(0,10),rep(1:9,each=50))+rnorm(460,sd=0.05),result,
       main="Phenotypes NAM10",xlab="Cross",ylab="Phenotype",
       ylim=c(min(result),max(result)+1))
  abline(v=c(0.5:8.5))
  text(c(0:10),rep(69.7,10),c("parents",paste0("A0xA",1:9)))
  dev.off()
  
  data<-data.frame(pop=c(rep("parent",10),rep(1:9,each=50)),pheno=result)
  boxes<-sapply(unique(data$pop),function(x) data$pheno[which(data$pop==x)])
  names(boxes)<-c("parents",paste0("A0xA",1:9))
    
  png("Storage/Plots/NAM10_Ancestral_cross.png",width=800,height=650)
  boxplot(boxes,main="Boxplot of ancestral crosses",xlab="Cross",ylab="Phenotype")
  dev.off()
  
  png("Storage/Plots/NAM10_Boxplot_cross.png",width=600,height=500)
  boxplot(pheno~pop,data=data,main="Boxplot NAM10 per cross",
          xlab="Cross",ylab="Phenotype")
  dev.off()
  
  
  
    
  color<-colorspace::sequential_hcl(460)
  color<-colorRampPalette(c("red","blue"))
  
  pcoplot<-data.frame(pheno=sort(result),Q[order(result),])
  
  png("Storage/Plots/NAM10_PCOplot.png")
  plot(pcoplot[,2],pcoplot[,3],col=color(460),main="PCO NAM10",pch=19)
  dev.off()

```

The results ofexperiment three (NAM3 and NAM10 multiallelism) are processed here
```{r,Experiment_3}
Exp3<-readRDS("Storage/Exp3.RDS")

mod<-c("ancestral","biallelic","parental")
div<-list("NAM3 h02"=seq(1,15,3),
          "NAM3 h05"=seq(2,15,3),
          "NAM3 h08"=seq(3,15,3),
          "NAM10 h02"=seq(31,45,3),
          "NAM10 h05"=seq(32,45,3),
          "NAM10 h08"=seq(31,45,3))
QT<-c(1,1,1,3,3,3)
for(pop in 1:length(div)){
  for(model in 1:3){
    png(paste0("Storage/Plots/MultiQTL/",names(div)[pop],"_",mod[model],".png"),
        width=800,height=600,res=115)
    comp.skyplot(Exp3[seq(model,12,3)],pos[-6500:-6910],
                 index=div[[pop]],
                 main=paste(names(div)[pop],mod[model]),
                 QTLs=pos[QTLs[[QT[pop]]]],xindex=c(-6500:-6910),h=c(120,260),
                 legend=c("L","Q","K","K+Q"),min=F)
    dev.off()
    
    pvals<-sapply(Exp3[seq(model,12,3)],function(x){
      y<-x[-6500:-6910,div[[pop]]]
      y<-sort(as.vector(y))
      return(y)
    })
    png(paste0("Storage/Plots/MultiQTL/QQ",names(div)[pop],"_",mod[model],".png"),
        width=800,height=800,res=115)
    comp.QQplot(pvals,main=paste(names(div)[pop],mod[model]),
                coltype = "divergent",h=c(120,260),
                legend=c("L","Q","K","K+Q"))
    dev.off()
  }

}
```

In here Exp4 data are processed
```{r}
Result<-readRDS("Storage/Exp4_allelism2 .RDS")

names<-c("LB","LQB","MB","MQB","LA","LQA","MA","MQA","LP","LQP","MP","MQP")


for(j in 1:3){
  param<-lapply(Result,function(cross) cross[[j]])

  for(i in 1:4){
    x<-sapply(param,function(model) model[i])
    
    y<-lapply(1:ncol(x[[1]]),function(k){
        sapply(x,function(type){
        type[,k]
      })
    })
    
    if(j==1){
      y<-lapply(1:24,function(i){
        new<-matrix(NA,ncol=8,nrow=6910)
        rownames(new)<-map$marker
        for(j in 1:8){
          new[names(y[[i]][[j]]),j]<-y[[i]][[j]]
        }
        colnames(new)<-names(y[[i]])
        return(new)
      })
    }
    
    names(y)<-colnames(x[[1]])
    assign(names[i+4*(j-1)],y)
  }
}

Exp4<-list(LB,LQB,MB,MQB,LA,LQA,MA,MQA,LP,LQP,MP,MQP)
rm(LB,LQB,MB,MQB,LA,LQA,MA,MQA,LP,LQP,MP,MQP)
saveRDS(Exp4,"Storage/Exp4_processed.RDS")
```

The processed data is then analyzed using a series of plots
```{r}
Exp4<-readRDS("Storage/Exp4_processed.RDS")
names(Exp4)<-names
dim(Exp4$LB$QTL1h0.2_1)

comp.skyplot(Exp4$LB,pos,index=1:2,min=F)


explist<-list(c("LB","LQB","MB","MQB"),
              c("LA","LQA","MA","MQA"),
              c("LP","LQP","MP","MQP"))
name<-c("Biallelic","Ancestral","Parental")
tit<-paste("QTL",1:2)
tit<-paste(tit,"h")
tit<-paste0(tit,rep(c("0.2","0.5","0.8"),each=2))
tit<-paste(sort(tit),rep(paste("config",1:4),each=6))
tit<-sort(tit)
for(j in 1:3){
  for(i in 1:12){
  compare<-lapply(Exp4[explist[[j]]],function(model) model[[i]])
  title<-paste0(name[j],tit[i])
  title2<-paste(name[j],substr(tit[i],7,19))
  
  #Skyline plots
  if(i<=12){QTL<-1349
  }else{QTL<-5056}
   png(paste0("Storage/Plots/Multiallelism/","Skyplot_",title,".png"),
       width = 800,height=600,res=115)
  comp.skyplot(compare,pos,QTLs=pos[QTL],
               main=title2,min = F,legend=c("L","Q","K","Q+K"))
  dev.off()

  
  #QQplots
  pvals<-sapply(compare,function(x){
    sort(unlist(x),decreasing=T)
  })

  png(paste0("Storage/Plots/Multiallelism/","QQplot_",title,".png"),
      height=800,width=800,res=115)
  comp.QQplot(pvals,main=title2,
              coltype = "divergent",lim=3)
  dev.off()
}
}
res<--log10(Exp4$MP$QTL1h0.5_4[,1])
plot(pos,res)
abline(v=pos[1349])


View(phenotypes)
View(found_alleles)
```

The effect of frequencies on the p-values of Exp4 is analyzed, and the distribution of allelic frequencies is also plotted.
```{r,freq_pvals}
parents<-link_NAM(crossfounder[i],parental=F)
freqs<-apply(parents,1,table)
freqs2<-lapply(freqs,sort,decreasing=T)
mosaic.list(freqs,h=60,
            xlab="cM",ylab="Frequencies",main="Unordered Allele Frequency")
mosaic.list(freqs2,x=pos,xlab="cM",ylab="Frequencies",main="Ordered Allele Frequency")


#Correlation test for each model and each phenotype
Res<-readRDS("Storage/Exp4_allelism2 .RDS")
var<-apply(parents,1,function(x) var(as.character(x)))#can also be defined without as.character
Freq.cor<-list()
for(cross in 1:length(Res)){
  test<-lapply(Res[[cross]],function(mod){
    sm<-sapply(mod,function(x){
      if(is.null(rownames(x))){rownames(x)<-names(var)}
      var.c<-var[match(names(var),rownames(x))]
      var.c<-na.omit(var.c)
      test<-apply(x,2,function(y){
        t<-cor.test(y,var.c)
        return(c(test=t$p.value))
      })
      return(test)
    })
  })
  
  Freq.cor[[cross]]<-test
}
```



